### 2025-10-03 misc/expo/additive.tex
Tutorial 4

\floattext4.5 \ninebf Note.
\ninepoint
I am compiling this set of expository notes primarily to solidify these fundamental results of
additive combinatorics in
my own mind. Sources that are far more comprehensive exist on the web, so I'm not sure
how useful this will be to anyone else. Currently, the bulk of these notes come from
a series of YouTube lectures of Timothy Gowers (given in early 2022), though I intend
in future to add more relevant topics as I learn about them.

\bigskip

@@

This statement is called Freiman's theorem, as it was first proved by Freiman over the integers.
In these notes we will prove an analogue of it for subsets of $\FF_p^N$, in which ``$d$-dimensional arithmetic
progression'' is replaced by ``subspace''. First we need a lemma.

@@

We can flip the roles of $\delta$ and $N$ to arrive at the following statement,
equivalent to what we just proved: {\sl Any subset of $[N]$
of size at least $300/\log\log N$ contains a nontrivial arithmetic progression of length $3$.}

>>>>

\floattext4.5 \ninebf Note.
\ninepoint
I am compiling this set of expository notes primarily to solidify these fundamental results of
additive combinatorics in
my own mind. Sources that are far more comprehensive exist on the web, so I'm not sure
how useful this will be to anyone else. Currently, the bulk of these notes come from
a series of YouTube lectures of Timothy Gowers (given in early 2022), though I'm slowly adding more
relevant topics as I learn about them.
\vskip36pt
\noindent{\ninessbx I. Sumsets and the structure of small doubling}
\smallskip
\contline{1}{Pl\"unnecke's theorem}{1}
\smallskip
\contline{2}{The Balog--Szemer\'edi--Gowers theorem}{5}
\smallskip
\contline{3}{Freiman homomorphisms and Ruzsa embedding lemmas}{9}
\smallskip
\contline{4}{Rudiments of discrete Fourier analysis}{12}
\smallskip
\contline{5}{Bohr sets and Bogolyubov's lemma}{17}
\smallskip
\contline{6}{Freiman's theorem in vector spaces over fields of prime order}{19}
\medskip
\noindent{\ninessbx II. Arithmetic progressions}
\smallskip
\contline{7}{Roth's theorem}{22}
\medskip
\noindent{\ninessbx III. Miscellaneous theorems}
\smallskip
\contline{8}{Khovanskii's theorem}{27}
\smallskip
\contline{9}{The Furstenberg--S\'ark\"ozy theorem}{29}
\vskip36pt

@@

This statement is called Freiman's theorem, as it was first proved by Freiman over the integers.
A set $A$ for which the ratio
$|A+A|/|A|$ is small is said to have {\it small doubling}, so Freiman's theorem imposes structure on sets
with small doubling.
In these notes we will prove an analogue of it for subsets of $\FF_p^N$, in which ``$d$-dimensional arithmetic
progression'' is replaced by ``subspace''. First we need a lemma.

@@

We can flip the roles of $\delta$ and $N$ to arrive at the following statement,
equivalent to what we just proved: {\sl Any subset of $[N]$
of size at least $300N/\log\log N$ contains a nontrivial arithmetic progression of length $3$.}
